---
title: "Final Project"
author: "Philip Elgaard Brett"
date: "2022-12-08"
output: html_document
---

#Installing the necessary packages 
```{r}
library(tidyverse)
library(ggplot2)
```

#Importing my datasets into R
```{r}
BNP1 <- read_csv2("Data2/BNP_development.csv")
Voting3 <- read_csv2("Data2/Voting.csv")
```

##Inspecting my data 
Using the str() function it can get an overview of how RStudio reads it. 
```{r}
str(Voting3)
```
I can see that it reads it correctly, with both Year and Percentage of Votes being registered as numeric values, and Party being read as characters

#Plotting my GDP data 
I'm plotting my data into a graph to show the GDP (BNP in danish) development in Denmark from 1970-2021
Using geom_col() to show the development. 
Also using the scale_x_continuous function I found on https://www.statology.org/ggplot2-scale_x_continuous/ to make the visualization more precise. 
```{r}
ggplot(data = BNP1) + 
  geom_col(aes(x = Year, y = BNP)) +
  scale_y_continuous(breaks = seq(-6, 6, by = 1)) +
  scale_x_continuous(breaks = seq(1970, 2022, by = 4)) +
  labs(title = 'BNP development')
```

##Trying another visualization 
To get another overlook I plot my data again, this time using geom_line()
```{r}
ggplot(data = BNP1) + 
  geom_line(aes(x = Year, y = BNP, color = 'Red')) +
    scale_y_continuous(breaks = seq(-6, 6, by = 1)) +
    scale_x_continuous(breaks = seq(1970, 2022, by = 4)) ##Husk at kildehenvise koden 
```

#Plotting overall election results
Now I am plotting my data on election results from 1971-2019 using geom_col() and adding color. 
```{r}
ggplot(data = Voting3) +
  geom_col(aes(x = Year, y = Percentage_of_votes, color = Party, fill=Party)) + 
  labs(title = "Voting Percentage")
```
#Plotting overall election results again 
To get a clearer and more pleasing visualization, I use geom_line() this time
```{r}
ggplot(data = Voting3) +
  geom_line(mapping = aes(x = Year, y = Percentage_of_votes, color = Party)) +
  labs(title = "Voting")
```
##Merging the parties into groups 
The visualization above is a bit messy, because of all the different parties, so in order to make the visualizations simpler and more transparent, I will merge the different parties into three groups. Left-wing parties, right-wing parties and center parties. '

I will use the filter() function so merge the different groups 
```{r}
Left_wing <- Voting3 %>% 
  filter(Party == "SF" | Party == "EL" | Party == "ALT" | Party == "VS" | Party == "DKP" | Party == "FK")
```

Now visualizing the data left-wing data with geom_line 
```{r}
ggplot(data = Left_wing) +
  geom_line(mapping = aes(x = Year, y = Percentage_of_votes, color = Party)) +
  labs(title = "Left wing votes 1971-2019")
```
Now doing the same with right-wing parties 
```{r}
Right_wing <- Voting3 %>% 
  filter(Party == "DF" | Party == "FP" | Party == "NB" | Party == "LA")
```

Visualizing the right-wing votes 
```{r}
ggplot(data = Right_wing) +
  geom_line(mapping = aes(x = Year, y = Percentage_of_votes, color = Party)) +
  labs(title = "Right wing votes 1971-2019")
```
Merging the danish danish centre parties 
```{r}
Centre_parties <- Voting3 %>% 
  filter(Party == "S" | Party == "V" | Party == "KF" | Party == "RV" | Party == "CD" | Party == "DR" | Party == "KRF/KD")
```

Visualizing the center-parties 
```{r}
ggplot(data = Centre_parties) +
  geom_line(mapping = aes(x = Year, y = Percentage_of_votes, color = Party)) +
  labs(title = "Centre votes 1971-2019")
```

##Summarizing my Left, Center and Right-wing votes in order to make the final visualisation more straightforward
Summarizing my center parties 
```{r}
Centre_parties_summarized <- Centre_parties %>% 
  drop_na(Percentage_of_votes) %>% 
  group_by(Year) %>% 
  summarise(total_block_votes_pct = sum(Percentage_of_votes)) 
```

Summarizing the Left Wing votes into one group.
```{r}
Left_parties_summarized <- Left_wing %>%
  drop_na(Percentage_of_votes) %>% 
  group_by(Year) %>% 
  summarise(total_block_votes_pct = sum(Percentage_of_votes))
```

Summarizing the Right Wing votes into one gruop 
```{r}
Right_parties_summarized <- Right_wing %>% 
  drop_na(Percentage_of_votes) %>% 
  group_by(Year) %>% 
  summarise(total_block_votes_pct = sum(Percentage_of_votes))
```

##Visualizating the three groups 
Now showing the three groups together with geom_line()
```{r}
ggplot() + 
   geom_line(data = Left_parties_summarized, aes(x = Year, y = total_block_votes_pct, color = "Left-wing")) + 
  geom_line(data = Right_parties_summarized, aes(x = Year, y = total_block_votes_pct, color = "Right -wing")) + 
  geom_line(data = Centre_parties_summarized, aes( x= Year, y = total_block_votes_pct, color = "Centre-parties")) +
  scale_x_continuous(breaks = seq(1970, 2022, by = 4))
```
##Visualizing GDP development and voting trends together 
Applying multiple dataframes to show GDP development and election results together and making them easier to compare. 
```{r}
ggplot() + 
  geom_line(data = Left_parties_summarized, aes(x = Year, y = total_block_votes_pct, color = "Left-wing")) + 
  geom_line(data = Right_parties_summarized, aes(x = Year, y = total_block_votes_pct, color = "Right -wing")) + 
  geom_line(data = Centre_parties_summarized, aes( x= Year, y = total_block_votes_pct, color = "Centre-parties")) + 
  geom_line(data = BNP1, aes(x = Year, y = BNP1$BNP, color = "BNP")) +
  scale_x_continuous(breaks = seq(1970, 2022, by = 4))
```
#Conclusion 
I can now see if there is any correlation between the fluctuation in the GDP development and election results. 
