---
title: "Web scraping W48"
author: "Philip Elgaard Brett"
date: "2022-12-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Task 
I am doing task 2.2) and using the 'rvest' library to scrape data of my interest. I will be using worldometers data on world population do determine the yearly change in population in the largest countries in the world. 

##Solution 
First I install the neccesary R packages 

```{r libraries}
library(rvest)
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```

##Scraping the data
Next, I will scrape the content of the website and extract the HTML table:

```{r scrape-cells}
url <- "https://www.worldometers.info/world-population/population-by-country" 
  url_html <- read_html(url) 
```

##I now extract the whole HTML table through the <table> tag.

```{r scrape-table}
population <- url_html %>% 
 html_nodes("table") %>%
 html_table()  #str(population) turns out to be a list
str(population)
population[[1]]
population
```

##Converting 
Now I convert my webscraping to a useful dataframe in R

```{r html-to-df}
population<- do.call(cbind,unlist(population, recursive = FALSE)) 
head(population)
```
I am now done with the scraping and have the data from the database sheet in a dataframe.

I have a clean table so I can see the yearly change in population in the largest countries. 